---
title: "NYPD Stop-Question-Frisk Part 4 Association Rules Mining"
author: "zehengl"
date: "January 27, 2016"
output: html_document
---

# Load the Cleansed Dataset
```{r, cache=TRUE}
load("SQF_clean.rda")
```

# Convert dataset to transactions
In order to perform itemsets mining we need to convert our dataset to transactions
```{r}
library(arules)
print( try(trans <- as(dat, "transactions")) )
colnames(dat)[c(2, 3, 6, 7, 74, 79, 80)]
```
As suggested by the error message, we need to discretize those columns. "xcoord" and "ycoord" are not removed.

```{r}
datestop <- dat[,"datestop"]
dat[,"datestop"] <- discretize(datestop,method="interval",categories=12)

timestop <- dat[,"timestop"]
dat[,"timestop"] <- discretize(timestop,method="interval",categories=24)

perobs <- dat[,"perobs"]
dat[,"perobs"] <- discretize(perobs,method="frequency",categories= 3)

perstop <- dat[,"perstop"]
dat[,"perstop"] <- discretize(perstop,method="frequency",categories= 3)

age <- dat[,"age"]
dat[,"age"] <- discretize(timestop,method="frequency",categories= 3)

dat[,"xcoord"] <- NULL
dat[,"ycoord"] <- NULL
```

Let's create a new column to represent whether a pedestrian is armed and clean out some other columns.
```{r}
dat$armed <- dat$pistol | dat$riflshot | dat$asltweap | dat$knifcuti | 
  dat$machgun | dat$othrweap | dat$contrabn
for (i in c("contrabn","pistol","riflshot","asltweap","knifcuti","machgun","othrweap")){
  dat[,i]<-NULL
}

for (i in c("offverb","offshld","officrid","offunif","typeofid")){
  dat[,i]<-NULL
}
```

Now Let's redo the transaction conversion.
```{r}
print( try(trans <- as(dat, "transactions")) )
```

```{r}
itemFrequencyPlot(trans,topN=20)

plot(sort(itemFrequency(trans, type="absolute"), decreasing=TRUE),
     xlab = "Items", ylab="Support Count", type="l")
```

```{r}
freqs <- apriori(trans, parameter=list(target="frequent", support=0.5, minlen=2, maxlen=5))
freqs <- sort(freqs, by="support")
inspect(head(freqs, n=10))
rules <- apriori(trans, parameter=list(target="rules", support=0.5, minlen=2, maxlen=5))
rules <- sort(rules, by="lift")
inspect(head(rules, n=10))

```

