---
title: "NYPD Stop-Question-Frisk Part 1 Data Cleansing" 
author: "zehengl"
date: "January 18, 2016"
output: html_document
---

The Stop, Question and Frisk Data from NYPD is available at http://www.nyc.gov/html/nypd/html/analysis_and_planning/stop_question_and_frisk_report.shtml

Download the 2014 archive and specs, unzip, and let the party begins.

# Getting Familiar with Data

```{r, cache=TRUE}
dat <- read.csv("2014.csv")
```
First, load 2014 SQF dataset and take a look at the data. There are `r dim(dat)[1]` rows and `r dim(dat)[2]` columns.

```{r}
head(dat)
```

In order to comprehend the information captured in the csv file, you need to go through the specs document. For example, column "pct" stands for precinct of stop; column "perobs" stands for period of observation in minutes; etc. 

# Cleasing the data

```{r}
# 1 YEAR OF STOP (CCYY)
dat$year <- NULL

# 2 PRECINCT OF STOP (FROM 1 TO 123)
dat$pct <- as.factor(dat$pct)

# 3 UF250 SERIAL NUMBER
dat$ser_num <- NULL

# 4 DATE OF STOP (MM-DD-YYYY)
dat$datestop <- as.Date(sprintf("%08d", dat$datestop), format ="%m%d%Y")

# 5 TIME OF STOP (HH:MM)
dat$timestop <- as.integer(substr(sprintf("%04d", dat$timestop), 1, 2))

# 6 RECORD STATUS
dat$recstat <- NULL

# 7 WAS STOP INSIDE OR OUTSIDE ?
dat$inout <- as.factor(dat$inout)

# 8 WAS LOCATION HOUSING OR TRANSIT AUTHORITY ?
dat$trhsloc <- as.factor(dat$trhsloc)
```

```{r}
# 9 PERIOD OF OBSERVATION (MMM)
summary(dat$perobs)
```
Clearly there are some outliers. Let's remove those greater than 2 hours (120).

```{r}
# 9 PERIOD OF OBSERVATION (MMM)
dat$perobs[ dat$perobs>120 ] <- NA
dat$perobs <- as.integer(as.character(dat$perobs))

# 10 CRIME SUSPECTED
dat$crimsusp <- NULL
```

```{r}
# 11 PERIOD OF STOP (MMM)
summary(dat$perstop)
```
"**" is not a valid input. Let's remove it.

```{r}
# 11 PERIOD OF STOP (MMM)
dat$perstop[dat$perstop =="**"] <- NA
dat$perstop <- as.integer(as.character(dat$perstop))

# 12 STOPPED PERSON'S IDENTIFICATION TYPE
dat$typeofid <- as.factor(dat$typeofid)

# 13 DID OFFICER EXPLAIN REASON FOR STOP ?
dat$explnstp <- dat$explnstp=="Y"

# 14 WERE OTHER PERSONS STOPPED, QUESTIONED OR FRISKED ?
dat$othpers <- dat$othpers=="Y"

# 15 WAS AN ARREST MADE ?
dat$arstmade <- dat$arstmade=="Y"

# 16 OFFENSE SUSPECT ARRESTED FOR
dat$arstoffn <- NULL

# 17 WAS A SUMMONS ISSUED ?
dat$sumissue <- dat$sumissue=="Y"

# 18 OFFENSE SUSPECT WAS SUMMONSED FOR
dat$sumoffen <- NULL

# 19 COMPLAINT YEAR (IF COMPLAINT REPORT PREPARED)
dat$compyear <- NULL

# 20 COMPLAINT PRECINCT (IF COMPLAINT REPORT PREPARED)
dat$comppct <- NULL

# 21 WAS OFFICER IN UNIFORM ?
dat$offunif <- dat$offunif=="Y"

# 22 ID CARD PROVIDED BY OFFICER (IF NOT IN UNIFORM)
dat$officrid <- NULL

# 23 WAS SUSPECT FRISKED ?
dat$frisked <- dat$frisked=="Y"

# 24 WAS SUSPECT SEARCHED ?
dat$searched <- dat$searched=="Y"

# WAS ARMED?
  # 25 WAS CONTRABAND FOUND ON SUSPECT ?
  # 27 WAS A PISTOL FOUND ON SUSPECT ?
  # 28 WAS A RIFLE FOUND ON SUSPECT ?
  # 29 WAS AN ASSAULT WEAPON FOUND ON SUSPECT ?
  # 30 WAS A KNIFE OR CUTTING INSTRUMENT FOUND ON SUSPECT ?
  # 31 WAS A MACHINE GUN FOUND ON SUSPECT ?
  # 32 WAS ANOTHER TYPE OF WEAPON FOUND ON SUSPECT
binary <- strsplit("contrabn pistol riflshot asltweap knifcuti machgun othrweap",   " ")[[1]]
for(b in binary) dat[[b]] <- dat[[b]]=="Y"

# 26 WERE ADDITIONAL REPORTS PREPARED ?
dat$adtlrept <- dat$adtlrept=="Y"

# PHYSICAL FORCE USED BY OFFICER
  # 33 PHYSICAL FORCE USED BY OFFICER - HANDS
  # 34 PHYSICAL FORCE USED BY OFFICER - SUSPECT AGAINST WALL
  # 35 PHYSICAL FORCE USED BY OFFICER - SUSPECT ON GROUND
  # 36 PHYSICAL FORCE USED BY OFFICER - WEAPON DRAWN
  # 37 PHYSICAL FORCE USED BY OFFICER - WEAPON POINTED
  # 38 PHYSICAL FORCE USED BY OFFICER - BATON
  # 39 PHYSICAL FORCE USED BY OFFICER - HANDCUFFS
  # 40 PHYSICAL FORCE USED BY OFFICER - PEPPER SPRAY
  # 41 PHYSICAL FORCE USED BY OFFICER - OTHER
for(b in grep("pf_", colnames(dat), value=TRUE)) dat[[b]] <- dat[[b]]=="Y"

# 42 RADIO RUN
dat$radio <- dat$radio=="Y"

# ADDITIONAL CIRCUMSTANCES
  # 43 ADDITIONAL CIRCUMSTANCES - REPORT BY VICTIM/WITNESS/OFFICER
  # 44 ADDITIONAL CIRCUMSTANCES - ONGOING INVESTIGATION
  # 47 ADDITIONAL CIRCUMSTANCES - PROXIMITY TO SCENE OF OFFENSE
  # 56 ADDITIONAL CIRCUMSTANCES - EVASIVE RESPONSE TO QUESTIONING
  # 57 ADDITIONAL CIRCUMSTANCES - ASSOCIATING WITH KNOWN CRIMINALS
  # 60 ADDITIONAL CIRCUMSTANCES - CHANGE DIRECTION AT SIGHT OF OFFICER
  # 65 ADDITIONAL CIRCUMSTANCES - AREA HAS HIGH CRIME INCIDENCE
  # 68 ADDITIONAL CIRCUMSTANCES - SIGHTS OR SOUNDS OF CRIMINAL ACTIVITY
  # 69 ADDITIONAL CIRCUMSTANCES - OTHER
  # 66 ADDITIONAL CIRCUMSTANCES - TIME OF DAY FITS CRIME INCIDENCE
for(b in grep("ac_", colnames(dat), value=TRUE)) dat[[b]] <- dat[[b]]=="Y"

# REASON FOR FRISK
  # 45 REASON FOR FRISK - VIOLENT CRIME SUSPECTED
  # 46 REASON FOR FRISK - OTHER SUSPICION OF WEAPONS
  # 48 REASON FOR FRISK - INAPPROPRIATE ATTIRE FOR SEASON
  # 53 REASON FOR FRISK-  ACTIONS OF ENGAGING IN A VIOLENT CRIME
  # 59 REASON FOR FRISK - REFUSE TO COMPLY W OFFICER'S DIRECTIONS
  # 61 REASON FOR FRISK - VERBAL THREATS BY SUSPECT
  # 67 REASON FOR FRISK - KNOWLEDGE OF SUSPECT'S PRIOR CRIM BEHAV
  # 76 REASON FOR FRISK - FURTIVE MOVEMENTS
  # 77 REASON FOR FRISK - SUSPICIOUS BULGE
for(b in grep("rf_", colnames(dat), value=TRUE)) dat[[b]] <- dat[[b]]=="Y"

# REASON FOR STOP
  # 49 REASON FOR STOP - CARRYING SUSPICIOUS OBJECT
  # 50 REASON FOR STOP - FITS A RELEVANT DESCRIPTION
  # 51 REASON FOR STOP - CASING A VICTIM OR LOCATION
  # 52 REASON FOR STOP - SUSPECT ACTING AS A LOOKOUT
  # 54 REASON FOR STOP - WEARING CLOTHES COMMONLY USED IN A CRIME
  # 55 REASON FOR STOP - ACTIONS INDICATIVE OF A DRUG TRANSACTION
  # 58 REASON FOR STOP - FURTIVE MOVEMENTS
  # 62 REASON FOR STOP - ACTIONS OF ENGAGING IN A VIOLENT CRIME
  # 63 REASON FOR STOP - SUSPICIOUS BULGE
  # 64 REASON FOR STOP - OTHER
for(b in grep("cs_", colnames(dat), value=TRUE)) dat[[b]] <- dat[[b]]=="Y"

# BASIS OF SEARCH
  # 70 BASIS OF SEARCH - HARD OBJECT
  # 71 BASIS OF SEARCH - OUTLINE OF WEAPON
  # 72 BASIS OF SEARCH - ADMISSION BY SUSPECT
  # 73 BASIS OF SEARCH - OTHER
for(b in grep("sb_", colnames(dat), value=TRUE)) dat[[b]] <- dat[[b]]=="Y"

# 74 REPORTING OFFICER'S COMMAND (1 TO 999)
dat$repcmd <- as.factor(dat$repcmd)

# 75 REVIEWING OFFICER'S COMMAND (1 TO 999)
dat$revcmd <- as.factor(dat$revcmd)

# 78 VERBAL STATEMENT PROVIDED BY OFFICER (IF NOT IN UNIFORM)
dat$offverb <- as.factor(dat$offverb)

# 79 SHIELD PROVIDED BY OFFICER (IF NOT IN UNIFORM)
dat$offshld <- as.factor(dat$offshld)

# 80 REASON FORCE USED
dat$forceuse <- as.factor(dat$forceuse)

# 81 SUSPECT'S SEX
dat$sex <- as.factor(dat$sex)

# 82 SUSPECT'S RACE
dat$race <- as.factor(dat$race)

# 83 SUSPECT'S DATE OF BIRTH (CCYY-MM-DD)
dat$dob <- NULL
```

```{r}
# 84 SUSPECT'S AGE
dat$age <- as.integer(as.character(dat$age))
summary(dat$age)
```
Clearly there are some outliers. Let's remove those less than 10 or more than 90.

```{r}
# 84 SUSPECT'S AGE
dat$age[ dat$age < 10 | dat$age > 90 ] <- NA
summary(dat$age)
# 85 SUSPECT'S HEIGHT (FEET)
dat$ht_feet <- NULL

# 86 SUSPECT'S HEIGHT (INCHES)
dat$ht_inch <- NULL

# 87 SUSPECT'S WEIGHT
dat$weight <- NULL

# 88 SUSPECT'S HAIRCOLOR
dat$haircolr <- as.factor(dat$haircolr)

# 89 SUSPECT'S EYE COLOR
dat$eyecolor <- as.factor(dat$eyecolor)

# 90 SUSPECT'S BUILD
dat$build <- as.factor(dat$build)

# 91 SUSPECT'S OTHER FEATURES (SCARS, TATOOS ETC.)
dat$othfeatr <- NULL

# 92 LOCATION OF STOP ADDRESS TYPE
# 93 LOCATION OF STOP RESIDENT CODE
# 94 LOCATION OF STOP PREMISE TYPE
# 95 LOCATION OF STOP PREMISE NAME
# 96 LOCATION OF STOP ADDRESS NUMBER
# 97 LOCATION OF STOP STREET NAME
# 98 LOCATION OF STOP INTERSECTION
# 99 LOCATION OF STOP CROSS STREET
# 100 LOCATION OF STOP APT NUMBER
# 102 LOCATION OF STOP STATE
# 103 LOCATION OF STOP ZIP CODE
# 104 LOCATION OF STOP ADDRESS PRECINCT
# 105 LOCATION OF STOP SECTOR
# 106 LOCATION OF STOP BEAT
# 107 LOCATION OF STOP POST
binary <- strsplit("addrtyp rescode premtype premname addrnum stname stinter crossst aptnum state zip addrpct sector beat post",   " ")[[1]]
for(b in binary) dat[[b]] <- NULL

# 101 LOCATION OF STOP CITY
dat$city <- as.factor(dat$city)

# 108 LOCATION OF STOP X COORD
dat$xcoord <- as.integer(dat$xcoord)

# 109 LOCATION OF STOP Y COORD
dat$ycoord <- as.integer(dat$ycoord)

# 110 DETAILS TYPES CODE
dat$dettypCM <- NULL

# 111 lineCM
dat$lineCM <- NULL

# 112 CRIME CODE DESCRIPTION
dat$detailCM <- as.factor(dat$detailCM)
```

After data cleansing, check the dimensions of new dataset.
```{r}
dim(dat)
```

Let's revisit the dataset.
```{r}
head(dat)
```

Now we save the cleansed dataset for future analysis.
```{r}
save(dat, file="SQF_clean.rda")
```